---
title: "RepAssignment1-Daily Activity"
author: "Okja Kim"
date: "November 18, 2016"
output: 
  html_document:
    keep_md: true
---

RepAssignment1
==============

## Daily Activity

###Loading and Processing the data

```{r}
unzip("activity.zip")
dataPA<- read.csv('Activity.csv', stringsAsFactors = FALSE)
dataPA$Date <- as.Date(dataPA$date)
```

###Mean and Median total number of steps taken per day
```{r}
library(ggplot2)
totalSteps<- tapply(dataPA$steps, dataPA$Date, FUN = sum, na.rm=TRUE)
qplot(totalSteps, binwidth=500, xlab = "Total number of steps taken per day")
```

Mean total number of steps taken per day:
```{r}
mean(totalSteps, na.rm = TRUE)
```

Median total number of steps taken per day:
```{r}
median(totalSteps, na.rm = TRUE)
```

###The average daily activity pattern
```{r}
averageSteps<- aggregate(x=list(steps=dataPA$steps), by=list(interval=dataPA$interval),
                         FUN=mean, na.rm=TRUE)
ggplot(averageSteps, aes(interval, steps))+geom_line()+xlab("5-minute interval")+ylab("average number of steps taken")
```

On average across all the days in the dataset, the 5-minute interval contains the maximum number of steps:
```{r}
averageSteps[which.max(averageSteps$steps),]
```


###Imputing missing values
There are a number of days/intervals where there are missing values.

```{r}
intervalsMissingValue<- is.na(dataPA$steps)
table(intervalsMissingValue)
length(which(intervalsMissingValue))
```
intervals where there are missing values.

Creating a new dataset that is equal to the original set but with the missing value filled in with the mean for that 5-minute intervals...
```{r}
newValue<-function(interval, steps){
    val<-steps
    if(is.na(val)) val<- averageSteps[averageSteps$interval==interval, "steps"]
    return(val)
}
newDataPA<- dataPA
newDataPA$steps<- mapply(newValue, newDataPA$interval, newDataPA$steps)
```

With the new dataset, histogram of the total number of steps taken
```{r}
tSteps<- tapply(newDataPA$steps, newDataPA$date, FUN = sum)
qplot(tSteps,binwidth=500, xlab = "Total number of steps taken each day")
```

Mean total number of steps taken per day:
```{r}
mean(tSteps, na.rm = TRUE)
```

Median total number of steps taken per day:
```{r}
median(tSteps, na.rm = TRUE)
```

The impact of imputting missing data on the estimates of the total daily number of steps is that the mean and median are higher than the original dataset.


###The differences in activity patterns between weekdays and weekends
```{r}
weekdayWeekend<- function(date){
  if (weekdays(date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
      return("weekday") else if (weekdays(date) %in% c("Saturday","Sunday"))
          return("weekend") else stop("invalid date")
}
newDataPA$day <- sapply(newDataPA$Date, FUN = weekdayWeekend)
```

Making a panel plot containing a time series plot of the 5-minute interval amd the average number of steps taken, averaged across all weekday days of weekend days...
```{r}
ave<- aggregate(steps~interval+day, data = newDataPA, mean)
ggplot(ave, aes(interval, steps))+geom_line()+facet_grid(day~.)+xlab("5-minute interval")+ylab("Number of steps")
```
